### 2023-05-29 2.State/main.tex

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The State:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{State of the art}
\label{cap:state}

\input{2.State/1.memristors.tex}
\input{2.State/2.crossbar.tex}
\input{2.State/3.model.tex}
\input{2.State/4.lstm.tex}

\cleardoublepage