### 2023-05-30 diapo/template1.tex

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{whale}

\usepackage{wrapfig}

\title[MSc Thesis]
{Memristors-based Recurrent Modules for Neural Computing}
%\subtitle{Using }

\institute[IST] % (optional)
{%
  Diogo Caetano\\
  INESC-MN
  \and%
  Ruxandra Barlulescu\\
  INESC-ID
}

\author[V. BARBAZA]{Valentin BARBAZA (MEEC)}

\date[2023] % (optional)
{May 2023}

\logo{\includegraphics[height=0.7cm]{figures/istnewlogo.pdf}}

\definecolor{istblue}{RGB}{0,102,153}
\setbeamercolor{titlelike}{bg=istblue}
\setbeamerfont{title}{series=\bfseries}

\begin{document}

\frame{\titlepage}

%\begin{frame}
%\frametitle{Table of Contents}
%\tableofcontents
%\end{frame}

\begin{frame}
\frametitle{Application/Motivation}
The goal of the thesis is to know whether an analog LSTM network would work. This would allow running a Neural Network in real-time in embedded systems. The analog implementation (using memristors crossbar array) allows for much greater computational speed and very small on-chip area.
\end{frame}
\begin{frame}
\frametitle{Technical Goals}
My point is to run simulations using cadence. Running a working simulation is the first step before fabricating the system.
\end{frame}
\begin{frame}
\frametitle{Progress}
So far, I have designed some smaller elements required for the system to work such as the activation function for example.\\
I've designed a full size LSTM (with 4 hidden state).
\begin{figure}
    \centering
    \includegraphics[height=0.35\textheight]{figures/LSTM.png}
\end{figure}
I am now working on a way to generate LSTM network of any size.
\end{frame}

\begin{frame}
\frametitle{What is missing}
I now need to run the simulations and compare the results of my simulation with the results a digital LSTM (running with Keras for example).
\end{frame}

\end{document}