### 2023-05-31 diapo/groupTemplate/main.tex
Changed a bit of the template, not done yet

\documentclass[table]{beamer}
\usetheme{Boadilla}
\usecolortheme{whale}

%\usepackage[table]{xcolor}

\title[Group Meeting]
{Weekly group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date[26 May] % (optional)
{26 May 2023}

\logo{\includegraphics[height=0.9cm]{figures/logos.png}}

\definecolor{text}{RGB}{25,116,146}
\definecolor{firstRow}{HTML}{219BC3}
\definecolor{evenRow}{HTML}{CCDEE9}
\definecolor{oddRow}{HTML}{E7EFF4}
\definecolor{istblue}{RGB}{0,102,153}
\setbeamercolor{titlelike}{bg=istblue}
\setbeamerfont{title}{series=\bfseries}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Introduction}
  Quick introduction to your topic.

  \begin{itemize}
      \color{text}
    \item Briefly summarized your project. Some of the attendants may need some refresh.
    \item Explain in which project stage you are, thus we all know how close you are to achieve your goals.
  \end{itemize}

  ex: I have to design a PCB to control a magnetic sensor. I have already done the PCB and now I am testing it.
\end{frame}

\begin{frame}
  \frametitle{Last weeks...}

  Last meeting, I promised to do…

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{3cm} c c c }
    \rowcolor{firstRow}
    & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
    1 & test2 & test23 & test3 & test4\\
    2 & test2 & test23 & test3 & test4\\
    3 & test2 & test23 & test3 & test4\\
    4 & test2 & test23 & test3 & test4\\
    5 & test2 & test23 & test3 & test4\\
  \end{tabular}

\end{frame}

\begin{frame}
  \frametitle{}
\end{frame}

\end{document}