### 2023-06-05 diapo/05-06-23_meeting/main.tex
Fixed logos

\documentclass[table]{beamer}
\usetheme{gpmeet}

\newcommand{\leftRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
\newcommand{\rightRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}

\usepackage{tikz}

\title{Group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date{05/06/2023}

\logo{
  \begin{tikzpicture}[overlay,remember picture]
    \node[left=0cm] at (current page.34){
      \includegraphics[height=1cm]{../figures/ist.eps}
      \hspace{1pt}
      \includegraphics[height=1cm]{figures/inesc-mn.png}
      \includegraphics[height=1cm]{../figures/inesc-id.eps}
    };
  \end{tikzpicture}
}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Introduction}

  \begin{itemize}
      \color{text}
    \item I am working on an analog implementation of a LSTM Neural Networks.
    \item The final objective of my thesis is to have a successful simulation of such a circuit.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \centering\includegraphics[width=\textwidth]{figures/LSTM-NP}
\end{frame}

\begin{frame}
  \frametitle{Last weeks...}

  Last meeting, I promised to do :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{3cm} c m{3cm} c }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
    1 & Try to use LTSpice & No & Found out I could do what I want with Cadence & No\\
    2 & Find another\newline problem to work on & No & Lack of time & Yes\\
    3 & Work on the thesis intro & on going &  & No\\
  \end{tabular}

\end{frame}

\begin{frame}{What's new}
  \begin{itemize}
    \item Results obtianed :
      \begin{itemize}
          \color{text}
        \item With the help of Fabio, I can now simulate all the steps automatically
        \item I successfully ran the simulation
        \item The result are not correct
      \end{itemize}
    \item Improvements :
      \begin{itemize}
          \color{text}
        \item Debug the whole thing
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{What's new}
  \includegraphics[width=\textwidth]{figures/fulltrial1-all-curves}
\end{frame}

\begin{frame}<1>[label=prob]
  \frametitle{Problems/doubts}
  \begin{itemize}
    \item<1-> The memory doesn't output the right values
    \item<2-> The memory input seems to be active for too long
  \item<3-> The output result isn't what is expected    \end{itemize}
\end{frame}

\begin{frame}{Problems/doubts}
  The memory doesn't output the right values.

  \centering
  \includegraphics[width=0.7\textwidth]{figures/fulltrial1-memory-prob}
\end{frame}

\againframe<2>{prob}

\begin{frame}{Problems/doubts}
  \centering
  \includegraphics[width=1\textwidth]{figures/fulltrial1-mem-act-late}
\end{frame}

\againframe<3>{prob}
\begin{frame}{Problems/doubts}
  The output is ~0.791 V. This is $-1.09$ and should be $0.015$

  \centering
  \includegraphics[width=1\textwidth]{figures/fulltrial1-final-res}
\end{frame}

\begin{frame}{Next weeks}

  List of what I plan to do in the next weeks :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{4cm} }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task \cr
    1 & Find another LSTM problem \\
    2 & "Debug" the circuit \\
    2.1 & Run the entire simulation \\
    3 & Work on thesis \\
  \end{tabular}
\end{frame}

\begin{frame}{Networking and sharing}
  \begin{tikzpicture}
    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}

    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
  \end{tikzpicture}
\end{frame}

\end{document}