### 2023-06-05 diapo/groupTemplate/main.tex

\documentclass[table]{beamer}
\usetheme{gpmeet}

\newcommand{\leftRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
\newcommand{\rightRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}

\usepackage{tikz}

\title{Group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date{26/05/2023}

\logo{
  \begin{tikzpicture}[overlay,remember picture]
    \node[left=0cm] at (current page.34){
      \includegraphics[height=1cm]{../figures/ist.eps}
      \hspace{1pt}
      \includegraphics[height=1cm]{figures/inesc-mn.png}
      \includegraphics[height=1cm]{../figures/inesc-id.eps}
    };
  \end{tikzpicture}
}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Introduction}
  Quick introduction to your topic.

  \begin{itemize}
      \color{text}
    \item Briefly summarized your project. Some of the attendants may need some refresh.
    \item Explain in which project stage you are, thus we all know how close you are to achieve your goals.
  \end{itemize}

  ex: I have to design a PCB to control a magnetic sensor. I have already done the PCB and now I am testing it.
\end{frame}

\begin{frame}
  \frametitle{Last weeks...}

  Last meeting, I promised to do…

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{3cm} c c c }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
    1 & test2 & test23 & test3 & test4\\
    2 & test2 & test23 & test3 & test4\\
    3 & test2 & test23 & test3 & test4\\
    4 & test2 & test23 & test3 & test4\\
    5 & test2 & test23 & test3 & test4\\
  \end{tabular}

\end{frame}

\begin{frame}{What's new}
  \begin{itemize}
    \item Results obtianed :
      \begin{itemize}
          \color{text}
        \item
      \end{itemize}
    \item Improvements :
      \begin{itemize}
          \color{text}
        \item
      \end{itemize}
    \item Relevant information :
      \begin{itemize}
          \color{text}
        \item test
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Problems/doubts}
  \begin{itemize}
    \item Things that must me sorted out in order to continue with your work.\\ Explain the problem.
    \item Things that you do not understand.
  \end{itemize}
\end{frame}

\begin{frame}{Next weeks}

  List of what I plan to do in the next weeks :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{4cm} }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task \cr
    1 & test23 \\
    2 & test23 \\
    3 & test23 \\
    4 & test23 \\
    5 & test23 \\
  \end{tabular}
\end{frame}

\begin{frame}{Networking and sharing}
  \begin{tikzpicture}
    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
  \end{tikzpicture}
\end{frame}

\end{document}