### 2023-06-26 diapo/26-06-23_meeting/main.tex
New group meeting

\documentclass[table]{beamer}
\usetheme{gpmeet}
\graphicspath{{../../figures/}}

\newcommand{\leftRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
\newcommand{\rightRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}

\usepackage{tikz}

\title{Group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date{12/06/2023}

\logo{
  \begin{tikzpicture}[overlay,remember picture]
    \node[below=1pt] at (current page.55){
      \includegraphics[height=1cm]{logos/ist}
      \hspace{1pt}
      \includegraphics[height=1cm]{logos/inesc-mn.png}
      \includegraphics[height=1cm]{logos/inesc-id.eps}
    };
  \end{tikzpicture}
}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Introduction}

  \begin{itemize}
      \color{text}
    \item I am working on an analog implementation of a LSTM Neural Networks.
    \item The final objective of my thesis is to have a successful simulation of such a circuit.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \centering\includegraphics[width=\textwidth]{lstm/lstm-np}
\end{frame}

\begin{frame}
  \frametitle{Last weeks...}

  Last meeting, I promised to do :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{3cm} c m{3cm} c }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
    1 & Try to use LTSpice & No & Found out I could do what I want with Cadence & No\\
    2 & Find another\newline problem to work on & No & Lack of time & Yes\\
    3 & Work on the thesis intro & on going &  & No\\
  \end{tabular}

\end{frame}

\begin{frame}{What's new}
  \begin{itemize}
    \item Results obtianed :
      \begin{itemize}
          \color{text}
        \item With the help of Fabio, I can now simulate all the steps automatically
        \item I successfully ran the simulation
        \item The result are not correct
      \end{itemize}
    \item Improvements :
      \begin{itemize}
          \color{text}
        \item Debug the whole thing
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{What's new}
  \includegraphics[width=\textwidth]{fulltrial1/fulltrial1-all-curves}
\end{frame}

\begin{frame}<1>[label=prob]
  \frametitle{Problems/doubts}
  \begin{itemize}
    \item<1-> The memory doesn't output the right values
    \item<2-> The memory input seems to be active for too long
    \item<3-> The output of the sigmoid doesn't look right (Now fixed)
    \item<4-> The output result isn't what is expected (Scale issue)
  \end{itemize}
\end{frame}

\begin{frame}{Problems/doubts}
  The memory doesn't output the right values.

  \centering
  \includegraphics[width=0.7\textwidth]{fulltrial1/fulltrial1-memory-prob}
\end{frame}

\againframe<2>{prob}

\begin{frame}{Problems/doubts}
  \centering
  \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-mem-act-late}
\end{frame}

\againframe<3>{prob}
\begin{frame}{Problems/doubts}
  Here we can see that the output of the sigmoid isn't right
  \begin{figure}[!tbp]
    \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{fulltrial1/sigmoid-prob.png}
    \end{minipage}
    \hspace{20pt}
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{activation/sigmoid-prob}
    \end{minipage}
  \end{figure}
\end{frame}

\begin{frame}{Problems/doubts}
  To fix that I had to re-adjust the reference voltages (sigmoid):
  \begin{figure}[!tbp]
    \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{activation/sigmoid}
      \caption{before}
    \end{minipage}
    \hspace{20pt}
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{activation/realSigmoid}
      \caption{after}
    \end{minipage}
  \end{figure}
\end{frame}

\begin{frame}{Problems/doubts}
  To fix that I had to re-adjust the reference voltages (tanh):
  \begin{figure}[!tbp]
    \centering
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{activation/tanh}
      \caption{before}
    \end{minipage}
    \hspace{20pt}
    \begin{minipage}[b]{0.4\textwidth}
      \centering
      \includegraphics[width=\textwidth]{activation/realTanh}
      \caption{after}
    \end{minipage}
  \end{figure}
\end{frame}

\againframe<4>{prob}
\begin{frame}{Problems/doubts}
  The output is weirdly on a different scale but gives out the right overall values. (Scale factor ~9022)

  \centering
  \includegraphics[width=.8\textwidth]{output/results0}
\end{frame}

\begin{frame}{Next weeks}

  List of what I plan to do in the next weeks :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{6cm} }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task \cr
    1 & Find another LSTM problem \\
    2 & Find out what is causing the issue \\
    3 & Find formula for approximate onChip area \\
    4 & Work on thesis \\
  \end{tabular}
\end{frame}

\begin{frame}{Networking and sharing}
  \begin{tikzpicture}
    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help\\- I would need the feature size of a typical memristor}

    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
  \end{tikzpicture}
\end{frame}

\end{document}