### 2023-07-04 2.State/main.tex
Changed state orga

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The State:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{State of the art}
\label{cap:state}

\input{2.State/nn.tex}
\input{2.State/lstm.tex}
\input{2.State/memristor.tex}
\input{2.State/crossbar.tex}
\input{2.State/model.tex}

\cleardoublepage