### 2023-07-07 3.Chapter/2.sectionb.tex
Moved placeholder figure

\section{Section B}
\label{sec:sectionb}

\subsection{Subsection A}
\label{subsec:subasectionB}

The model described can also be represented as

\begin{equation}
  \dot{\mathbf{x}}(t) = \mathbf{T}\mathbf{z}(y),\  \mathbf{y}(0) = \mathbf{y}_0,\  z\geq 0 \\
  \label{eq:dummyeq1}
\end{equation}

\noindent where

\begin{equation}
  \mathbf{A} = \left[ \begin{array}{cc} -(a_{12} + a_{10}) & a_{21} \\ a_{12} & -(a_{21} + a_{20}) \end{array} \right],\ \mathbf{x} = \left[ \begin{array}{c} x_1 \\ x_2 \end{array} \right] \\
    \label{eq:dummyeq2}
\end{equation}

\subsection{Subsection B}
\label{subsec:subbsectionB}

\begin{table}[H]
  \centering
  \caption{Dummy Table.}
  \begin{tabular}{|c|c|c|c|} \hline
    \textbf{Vendor Name}        & \textbf{Short Name} & \textbf{Commercial Name}  & \textbf{Manufacturer} \\ \hline \hline
    \multirow{3}{*}{Text in Multiple Row}   & ABC       &  ABC\textreg        & ABC SA               \\ \cline{2-4}
    &        DEF        &  DEF\textreg        & DEF SA        \\ \cline{2-4}
    &        GHF      &  GHF\textreg        & GHF SA        \\ \hline
    Text in Single Row          &        IJK        & IJK\textreg       & IJK SA        \\ \hline
    Frescos SA            &        LMN      & LMN\textreg       & LMN SA        \\ \hline
    Carros Lda.           &    \multicolumn{3}{|c|}{Text in Multiple Column}              \\ \hline
  \end{tabular}
  \label{tab:dummytable}
\end{table}

\begin{figure}[H]
  \centering
  \input{figures/TimeDiagram/basic.tex}
  \caption{Caption}
  \label{fig:my_label}
\end{figure}