### 2023-07-07 3.Circuit/activFunc.tex
Added chapter about circuits

\section{The activation functions}
\label{sec:af}

Activation function play a great role in the results obtained \cite{af}. It is the reason why making good Activation function circuits is fundamental. Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid. especially for regression problems \cite{hardSigm}.

\subsection{Subsection A}
\label{subsec:subasectionA}

This would be a citation \cite{dummy}.

\ac{acro}
% The first time you use this, the acronym will be written in full with the acronym in parentheses: supernova (SN). At later times it will just print the acronym: SN.

\acf{acro}
% written out form with acronym in parentheses, irrespective of previous use

\acs{acro}
% acronym form, irrespective of previous use

\acl{acro}
% written out form without following acronym

\acp{acro}
% plural form of acronym by adding an s. \acfp. \acsp, \aclp work as well.

As seen in \cite{wiki}. \emph{Enfatizar}

\subsection{Subsection B}
\label{subsec:subbsectiona}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{cover}
  \caption[Dummy Figure Caption for List of Figures.]{Dummy Figure Caption.}
  \label{fig:dummyfigure1}
\end{figure}

Remember you can change the reference style. Another dummy citation \cite{site}.