### 2023-07-07 3.Circuit/activFunc.tex
Removed duplicate example

\section{The activation functions}
\label{sec:af}

Activation function play a great role in the results obtained \cite{af}. It is the reason why making good Activation function circuits is fundamental. Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid. especially for regression problems \cite{hardSigm}.