### 2023-07-10 3.Circuit/activFunc.tex
Waiting for figure

\section{The activation functions}
\label{sec:af}

Activation functions play a great role in the results obtained \cite{af}. It is the reason why making good Activation function circuits is fundamental. Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.

Designing analog activation function as close to the original is very important for the final result's quality.

\subsection{The circuit)

  \subsection{The results}