### 2023-07-11 3.Design/main.tex
New figures plus renamed 3

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dummy Chapter:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The Introduction:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{The circuits}
\label{cap:circuit}

In this chapter, I'm going to describe the different circuit designed and used in this thesis.
%\textit{Present the chapter content.}

TODO : Maybe add more ???
Explain 0->0.9V and 1->1V!!

\input{3.Design/activFunc.tex}
\input{3.Design/1.sectiona.tex}
\input{3.Design/2.sectionb.tex}
\cleardoublepage