### 2023-07-12 3.Design/main.tex
change to the main text chap 3

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dummy Chapter:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The Introduction:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{The design}
\label{cap:design}

In this chapter, I'm going to describe the different design decisions taken during the study of the thesis.

The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.

The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$. That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$. This $V_{cm}$ will then describe a zero. A step of one was chosen to be $0.1V$ in the analog circuit.
Table \ref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.

\begin{table}[H]
\centering
\caption{Real/Voltage Conversion Table.}
\begin{tabular}{|c|c|}
\hline
\rowcolor{gray}
Real value & Voltage \\
       \hline
       0 & 0.9V \\
         \hline
         1 & 1.0V \\
           \hline
           $x$ & $\frac{x}{10}+V_{cm}$\\
             \hline
             $(V-0.9)\cdot 10$ & $V$\\
               \hline
               \end{tabular}
               \label{tab:valConv}
               \end{table}

               Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $V\in [0,1.8]$. This means that the range of real value that the systems can handle is $x\in [-9,9]$.

               Any data inputed in the analog system should have been previously checked to make sure it stays in the range of accepted values.

               \input{3.Design/activation.tex}
               \input{3.Design/1.sectiona.tex}
               \input{3.Design/2.sectionb.tex}
               \cleardoublepage