### 2023-07-14 2.State/lstm.tex
Small fix on figure + lstm state changes

\section{LSTM}\label{sec:lstm}
\acp{LSTM} are a type of \ac{NN} used to analyze sequence of data. They are capable of predict data based of the previous results. \acp{LSTM} are part of \acp{RNN}. \acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections. Meaning that the results from the last time step have an impact on the next time step.

The first \ac{LSTM} was invented in 1997 by Hochreiter and Schmidhuber \cite{firstLSTM}. \acp{LSTM} changed a lot through the years to become what they are now.
\acp{LSTM} fixed the vanishing/exploding gradient problem. Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation. This is achieved by connecting these sequential events with the hidden state:
$$h_t=f(h_{t-1},x_t)$$
The hidden state $h_t$ carries all the past informations for the next time step.
Even though \acp{RNN} work very well, they are subject to the vanishing/exploding gradient problem. This issue occurs when too many time steps are chained.

\acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This what gives \acp{LSTM} their use for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. (TODO : find example)

An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves. There is also a cell state which job is to hold a value for the next step.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{lstm/lstmCell.svg}
  \caption{LSTM Cell}
  \label{fig:lstmCell}
\end{figure}

Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a tanh or a sigmoid activation function. Each input to the cell is a vector.
Those vector are of varying sizes depending on several factors. For example, both $h_t$ and $c_t$ are of the same size as the number of hidden state (sometimes refered to as cell state) for any $t\geq 0$.
The input vector ($x_t$) is of size of the sample we want to feed for each time step.

\subsection{Equations}

The equations of an LSTM are quite unusual.
Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
The input (equation \ref{eq:inputG}), forget (equation \ref{eq:forgetG}) and ouput (equation \ref{eq:ouputG}) gates are described below.

\begin{equation}\label{eq:inputG}
  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i)
\end{equation}
\begin{equation}\label{eq:forgetG}
  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f)
\end{equation}
\begin{equation}\label{eq:ouputG}
  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o)
\end{equation}

Where ($w_i$,$b_i$), ($w_f$,$b_f$) and ($w_o$,$b_o$) are the pair of weights and biases for the input, forget and ouput gates respectively. $x_t$ is the input vector and $h_t$ is the hidden state vector.

For the next equation describes the candidate cell state (equation \ref{eq:candCell}), that will next become the cell state (equation \ref{eq:cellS}).

\begin{equation}\label{eq:candCell}
  \tilde{c}_t=tanh(w_c\cdot[x_{t-1},h_{t-1}] + b_c)
\end{equation}
\begin{equation}\label{eq:cellS}
  c_t=f_t\cdot c_{t-1} + i_t \cdot \tilde{c}_t
\end{equation}

Where $w_c$ and $b_c$ are the weights and bias for the candidate cell state.

The final step of the \ac{LSTM} is to compute the hidden state (equation \ref{eq:hiddenS}).
\begin{equation}\label{eq:hiddenS}
  h_t=o_t\cdot tanh(c_t)
\end{equation}

We set $x_0$ as the first input and define $h_0$ as a zero only vector.

\subsection{Usage}

Using \ac{LSTM} can be a bit tricky. Due to it's sequential nature, it takes several time steps. Every hidden state ($h_t$) is passed to the next time step as an input. This hidden state can be used to compute an output at any time step $t$.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{lstm/lstmUse}
  \caption{LSTM use case}
  \label{fig:lstmUse}
\end{figure}

\subsection{Variants}

\acp{LSTM} come in reality in a few different flavor. Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version. This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch. In fact is that there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
The difference between each of them varies from one to the other, but they can all be described with new equations.

\subsubsection{Vanilla}
This variant of the \ac{LSTM} was origanally the only \ac{LSTM} network, hence it's name. It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acf{NP}. The Vanilla \ac{LSTM} is thus defined by those equations \cite{vanillaLSTM, nbLSTM} :

\begin{equation}\label{eq:inputGVanilla}
  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i+c_{t-1}\odot p_f)
\end{equation}
\begin{equation}\label{eq:forgetGVanilla}
  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f+c_{t-1}\odot p_i)
\end{equation}
\begin{equation}\label{eq:ouputGVanilla}
  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o+c_{t}\odot p_o)
\end{equation}
With $p_f$, $p_i$ and $p_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($h_t$).

The equations that haven't been rewritten simply stay the same.