### 2023-09-06 2.State/model.tex
Tried to write on this part

\section{Memristor's model}\label{sec:model}
They are several ways to model a memristor (TODO : cite).

One of them, works by applying voltage pulse to the device. The device then change it's internal resistance based on how far from the stable point (usually the initial resistance) it is.
TODO