### 2023-09-06 3.Design/inverter.tex

\section{Inverter}
\label{sec:inv}

This section simply describes a very well know electrical component, the inverter. There is nothing special about this specific inverter. When it a voltage of $v_{dd}$ is applied to the input, the output is grounded and the same goes for the other way around.

\begin{figure}[H]
  \centering
  \hspace*{1.5cm}
  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[width=0.2\textwidth]{inverter/invCircuit}}%
  \hfill
  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{activation/tanhSymbol}}%
  \hspace*{1.5cm}
  \caption{}
  \label{fig:inv}
\end{figure}

Figure \ref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.