### 2023-09-07 Appendices/appendixA.tex
Added explication whyneed to amplify aft voltmult, and added appendix A, need to find name

\fancychapter{TODO determine name}
\label{ap:a}

\section{Amplificator}\label{apsec:amp}

This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it simply consists of two inverting amplifiers with one amplifying the results by 10. Figure TODO shows this simple circuit.

A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something