### 2023-09-08 Appendices/appendixA.tex

\fancychapter{TODO determine name}
\label{ap:a}

\section{Buffer}\label{apsec:buffer}

This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.

\section{Amplificator}\label{apsec:amp}

This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it simply consists of two inverting amplifiers with one amplifying the results by 10. Figure TODO shows this simple circuit.

A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something