### 2023-09-11 3.Design/inverter.tex
Worked on xbar + new package for refs + cleaning repo

\section{Inverter}
\label{sec:inv}

This section simply describes a very well know electrical component, the inverter. There is nothing special about this specific inverter. When it a voltage of $V_{dd}$ is applied to the input, the output is grounded and the same goes for the other way around.

\begin{figure}[H]
  \centering
  \hspace*{2.5cm}
  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[height=3cm]{inverter/invCircuit}}%
  \hfill
  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{inverter/invSymbol}}%
  \hspace*{1.5cm}
  \caption{}
  \label{fig:inv}
\end{figure}

\Cref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.

The approximate onChip area for the inverter is :
\begin{equation}
  A_{inv}=2\cdot A_{CMOS}
\end{equation}

With $A_{CMOS}= w\cdot l$ TODO check w et l