### 2023-09-12 2.State/crossbar.tex
crossbar advancement, plus changes to state crossbar for unity

\section{Memristors Crossbar Array}\label{sec:crossbar}

Setting memristors in a crossbar array allows to perform analog \ac{VMM} also called Multiply and Accumulate. This circuit works great for this use as all the computation is done almost instantly and using physics laws. Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.

\begin{figure}[H]
  \centering
  \subfloat[Schematics]{\includesvg[width=.45\linewidth]{crossbar/crossbar}}%\qquad
  \hfill
  \subfloat[3 dimensional representation]{\includegraphics[width=.45\linewidth]{crossbar/crossbar3D.jpg}}\\
  \caption{Memristor crossbar array}
  \label{fig:crossbar}
\end{figure}

It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
  \centering
  \includesvg[height=3.5cm]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$. As such and using Ohm's law, we know the the current flowing into the memristor ($I_{mem}$) is bound by the following equation :

\begin{equation}
  V_i = R\cdot I_{mem} \Rightarrow I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot\sigma_i
\end{equation}
With $\sigma_i$ being the conductance of memristor, defined as $\sigma_i=\frac{1}{R_i}$.

This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's current law the resulting current is :
\begin{equation}
  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot\sigma_i
\end{equation}
By applying this process to the whole system we find that the current as the bottom of one column is :
\begin{equation}
  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
\end{equation}
With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.

This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.%TODO elaborate