### 2023-09-12 3.Design/crossbar.tex

\section{Crossbar circuit}
\label{sec:xbarCircuit}

The crossbar circuit theory has already been explained in the \cref{sec:crossbar}. This section describes how the crossbar circuit is actually implemented. The final circuit is the one in \cref{circt:xbar}.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{crossbar/crossbarUse}
  \caption{Crossbar circuit used in the final system}
  \label{circt:xbar}
\end{figure}

\subsection{Two memristors per synapse}

First of all, a two memristors per synapse architecture has been used. This design is beneficial in two ways :

\begin{itemize}
  \item It doubles the range of weights (\cref{tab:synapses})
  \item It allows to easily use negative weights, as this design simply allows the output values to be centered around $V_{cm}$ to be compliant with the standard set in \cref{tab:valConv}.
\end{itemize}

This design is the one that is used in \cite{doubleMem}. Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means it's conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$). This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array. \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works.

% For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.

\begin{equation}
  \label{eq:doubleMem0}
  V_{opAmp0}=-R\cdot i_+ \Leftrightarrow
\end{equation}
\begin{equation}
  \label{eq:doubleMem1}
  i_{R_f}=i_-+\frac{V_{opAmp0}}{R}=i_--i_+
\end{equation}
With $i_{opAmp}=0A$ because we assume an ideal \ac{opamp}.
\begin{equation}
  \label{eq:doubleMem2}
  V_{opAmp2}=V_{out}=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
\end{equation}
With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.

In the previous equations, $x_k$ is the voltage for the input line $k$.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|c|}
    \cline{2-3}
    \rowcolor{gray}
    \multicolumn{1}{c|}{\cellcolor[HTML]{FFFFFF}} & Two memristors per synapse & One memristor per synapse \\
    \hline
    Maximum weight & $\sigma_{max}-\sigma_{min}$ & $\sigma_{max} -\overline{\sigma}$\\
    \hline
    Minimum weight & $\sigma_{min}-\sigma_{max}$ & $\sigma_{min} -\overline{\sigma}$\\
    \hline
    Range & $2\cdot(\sigma_{max}-\sigma_{min})$&$\sigma_{max}-\sigma_{min}$\\
    \hline
  \end{tabular}
  \caption{Synaptic weights precision (extracted from \cite{doubleMem})}
  \label{tab:synapses}
\end{table}

\subsection{Serialization/Parallelization}

The idea using the system in a serial mode came from \cite{thesisRef}. The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required. It also, and mainly reduces the point wise part of the \ac{LSTM} circuit.

This serial system has great advantages but comes with a few down sides. Let's assume the number of hidden state ($n_h$) is the major changing element in the final system(TODO ref final onchip area).

\begin{itemize}
  \item When the system is serialized, the inference time increases with a factor of $O(n_h)$.
  \item When the system is in parallel, the circuit onChip area increases with a factor of $O(n_h)$.
\end{itemize}

However, in both cases, the overall onChip area increases with a factor of $O(n_h^2)$. That means that using the parallel system won't increase the onChip area by a lot as $n_h$ rises, because the $O(n_h)$ factor of the parallel version is absorbed as $O(n_h^2+n_h)=O(n_h^2)$.

That leads to the conclusion that the best thing to do is use a both serialized and parallel system, using the limiting inference time for our system to set the maximum serial size.

%TODO understand better \subsection{Sneak path problem}
\subsection{Usage}
Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistance doesn't need to change during this phase. The weights only change during the training phase, and the resistances are just a rendition of the weights (section TODO : ref weigths to res)