### 2023-09-12 3.Design/wei2res.tex
Added new file for weight to resistance

\section{Conversion from weights to resistance}
\label{sec:wei2res}

This chapter will contain all the information on the weight conversion to a resistance value.