### 2023-09-14 diapo/15-09-23_meeting/main.tex
Finished pres demain

\documentclass[table]{beamer}
\usetheme{gpmeet}
\graphicspath{{../../figures/}}

\newcommand{\leftRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
\newcommand{\rightRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}

\usepackage{tikz}

\title{Group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date{26/05/2023}

\logo{
  \begin{tikzpicture}[overlay,remember picture]
    \node[left=0cm] at (current page.34){
      \includegraphics[height=1cm]{logos/ist.eps}
      \hspace{1pt}
      \includegraphics[height=1cm]{logos/inesc-mn.png}
      \includegraphics[height=1cm]{logos/inesc-id.eps}
    };
  \end{tikzpicture}
  }

  \begin{document}

  \frame{\titlepage}

  \begin{frame}
    \frametitle{Introduction}

    \begin{itemize}
        \color{text}
      \item I am working on an analog implementation of a LSTM Neural Networks.
      \item The final objective of my thesis is to have a successful simulation of such a circuit.
    \end{itemize}

  \end{frame}

  \begin{frame}
    \frametitle{Introduction}
    \centering\includegraphics[width=\textwidth]{lstm/lstm-np}
  \end{frame}

  \begin{frame}
    \frametitle{Last weeks...}

    There was no meeting last week, and a lot happened but here is what I'm doing :
    \begin{itemize}
      \item Mainly focused on writing the thesis
      \item I also need to run a new simulation with a new problem
    \end{itemize}
  \end{frame}

  \begin{frame}{What's new}
    \begin{itemize}
      \item Results obtianed :
        \begin{itemize}
            \color{text}
          \item None yet
        \end{itemize}
      \item Improvements :
        \begin{itemize}
            \color{text}
          \item
        \end{itemize}
      \item Relevant information :
        \begin{itemize}
            \color{text}
          \item
        \end{itemize}
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Problems/doubts}
    \begin{itemize}
      \item What is the approximate onChip area of an opAmp
      \item What is the precision of a memristor when setting the resistance
      \item I feel like my state of the art is missing content...
      \item When would my thesis defense happen ?
      \item I'm afriad i might not have time to finish on time wiht enough content.
    \end{itemize}
  \end{frame}

  \begin{frame}{Next weeks}

    List of what I plan to do in the next weeks :

    \centering
    \rowcolors{2}{evenRow}{oddRow}
    \begin{tabular}{ c m{6cm} }
      \rowcolor{firstRow}
      \color{white}\textbf{\#} & \centering\color{white}Task \cr
      0 & Work on thesis \\
      1 & Figure out the new problem \\
      2 & Digital inference time \\
    \end{tabular}
  \end{frame}

  \begin{frame}{Networking and sharing}
    \begin{tikzpicture}
      \leftRect{Help/Collaboration :}{- I'm looking for a software to create nice looking electronic circuit.}

      \rightRect{Recommendations :}{None}
    \end{tikzpicture}
  \end{frame}

\end{document}