### 2023-09-19 3.Design/genwei.tex
New files for weight generation

\section{Weights generation}
\label{sec:weigen}

The choice for the weight generation was to use Keras using the tensorflow implementation. However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.

Generating the weights simply requires to train the \ac{NN} with the required achitecture. Once the trainning is done, we can run an inference of the \ac{NN} to check whether the training was good. Then the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved as a file. The files can now be taken to generate the netlist.

The code used for all the weights generations is available at \cite{lstmWei}.