### 2023-09-20 3.Design/fullsys.tex
Finishing at the library

\section{Full analog system}\label{sec:fullsys}

\subsection{Circuit controller}%TODO maybe move in annex further

The final circuit will need an external controller to manage diffenrent things in the circuit. This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex

The controller will increase the power consumption of the overall system. However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and won't affect the overall power consumption of the final system.

The controller will have several roles in the system :
\begin{itemize}
\item The first role will be to manage the different flags. This means that the controller will have to set those to high or low in order to control the system.
\end{itemize}