### 2023-09-21 3.Design/netlist.tex

\section{Netlist generation}
\label{sec:netlist}

In order to be able to run the simulation with different kinds of \ac{NN} architecture. The point of the part is thus to explain how the netlist generator tool works. The tool also work to generate any architecture with the supported layers (listed in the README.md of \cite{lstmGen}).

This is done by generating a SPICE netlist using a python script. This script takes in a few parameters :

\begin{itemize}
  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($x_t$).
  \item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
  \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network. This parameter must divide the number of hidden states in the \acp{LSTM} layers.
  \item The files containing different info about the model. This file contains the type of \ac{NN} architecture and the weights associated with each layer. The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
  \item Finally the name of the file in which the output of the script (the netlist) will be written to.
\end{itemize}

\subsection{Weights storage}\label{subsec:weiStore}

The weights are stored in a single file that is the save a numpy array. The first dimension contains the weights for the different layers. Depending on the layer used, the weights are stored a bit differently :

\begin{itemize}
  \item Dense layer : This is basically a \ac{VMM} so the weights are just sorted linearly in a list. The list is of size $n^2$, where $n$ is the size of the input vector. \Cref{mtrx:wei} shows the position ($i$) of the weight ($w_i$) in the matrix.
  \item \ac{LSTM} layer : An \ac{LSTM} layer contains four \acp{VMM} which can be assimilated and thus stored like a dense layer. The weights for each \ac{VMM} is then stored in a sub list.
\end{itemize}

\begin{equation}\label{mtrx:wei}
  \begin{bmatrix}
    w_{0} & w_{1} & \dots \\
    \vdots & w_i & \vdots \\
    \dots & w_{n^2-2} & w_{n^2-1}\\
  \end{bmatrix}
\end{equation}

All the code can be found on my github page \cite{lstmGen}.