### 2023-09-22 Appendices/appendixA.tex

\fancychapter{TODO determine name}
\label{ap:a}

\section{Buffer}\label{apsec:buffer}

This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.

\section{Inverting amplifier}\label{apsec:invAmp}

This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10. Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).

A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something