### 2023-09-25 3.Design/lstm.tex
Trying to add the lstm circuit figure

\section{\ac{LSTM} analog implementation}\label{sec:lstmCircuit}

\subsection{Inputs}%TODO oen section about datasets ?

\begin{figure}[H]
\centering
\includesvg[width=\textwidth]{lstm/lstmCircuit.svg}
\caption{\ac{LSTM} circuit}
\label{circt:lstm}
\end{figure}