### 2023-10-05 1.Introduction/motivation.tex
Improved motivation

\section{Motivation}
\label{sec:int_motivation}

Nowadays, it's impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them. The most recent example being \textit{ChatGPT} from \textbf{OpenAI} and is getting very popular even among casual computer users. Everybody is using those tools. It's clear that \ac{AI} is becoming more and more important in the current world.

\acp{AI} are very power hungry algorithms which limits its use in embedded systems and power efficient devices.
Furthermore, the time an \ac{AI} takes to compute its output quickly rises, especially on lower end chip such as the ones used in embedded systems. So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT}.

There are several options to reduce execution speed and energy consumption such as running the algorithm on a \ac{GPU} rather than on a \ac{CPU}. Using an \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
The latter is the most restricting (as it's name implies) but has the best results.

Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it's much faster while being very power-efficient.

Making an analog chip with a very low power consumption allows to use chip like that in embedded systems such as video surveillance cameras. A small chip could be installed inside the system that could be able to do gait detection \cite{gaitDS,gaitDig,gait}. The video surveillance camera would then be able to know who is showing up on the camera and if it's not a known person send an alert.

Such a chip could also be used for video stabilization \cite{videoStab}. It would allow to directly store stabilized video, and not have to use online servers like described in \cite{videoStab}. Any action camera would benefit greatly from such an improvement since they're mostly used with a lot of movements.

Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}. That is useful especially for VR headsets that need to know their location in the room. The embedded chip would allow this computation to be much faster than other methods, and thus give more realistic results.

Somre research \cite{celegans}, focuses on reproducing parts of the nervous system of a small organism. This could be applied on larger organisms, or, in the long term, reproduce part of the human brain. Using analog computers, small chips could be used to replace a part of one's brain non functionnal part.

The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's virtuoso as simulation software. Having those preliminary results would allow for fabricating and testing such a chip. Once a working prototype is manufactured, such a circuit could highly improve embedded \ac{AI} use.