### 2023-10-10 3.Design/dataset.tex
Dataset chap update

\section{The datasets}\label{sec:dataset}

\subsection{The format}

The values for the data needs to be set not to exceed 1. This is to make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.

\subsection{Airline}

The first dataset contains a time series of international airline passengers from January 1949 to December 1960, the data is recorded monthly and is in thousands. The dataset then contains twelve years of monthly data so $144$ sample points. The results of such a problem aren't useful, it was used in this thesis because of its simplicity to check whether the analog circuit is working. It is extracted from the tutorial available at \cite{airline}. The dataset is available at \cite{datasets}, the name of the file containing the dataset is \textit{airline-passengers.csv}.

This is a regression problem, the role of the \ac{LSTM} is to predict the number of passenger flying the next month being given previous months' passenger count.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{datasets/airline}
  \caption{the airline dataset in a grphic. This is the curve that the regression is trying to reproduce. The vertical lines represent a full year.}
  \label{graph:airline}
\end{figure}

\subsubsection{Data format}

The dataset contains $144$ sample points. It has been transformed into $142$ data points for training. This has been done by taking each values three by three. The first two values are two timesteps for the input vector and the third value being used the target value. Two third of the dataset is being used for training and the other third is used for validation.

\subsubsection{Network configuration}

The layers used to solve this problem are listed below :

\begin{itemize}
  \item An \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
  \item A Dense layer with an output size of one.
\end{itemize}

\Cref{fig:airlineModel} is a graphical representation of the model just described.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{datasets/airlineModel}
  \caption{Model used to solve the airline passengers problem}
  \label{fig:airlineModel}
\end{figure}

\subsubsection{Results} %TODO Move to results obviously

The weights are trained using both the custom activation function (\cref{sec:genwei}) and without ones and compare their output.

After training on a digital computer using the settings described in \cref{sec:genwei}.

\subsection{\ac{C. elegans}}

This dataset is far more interresting than the latter. This data set aims to use \acp{LSTM} to mimic the behavior of real neurons. As explained in \cite{celegans}, \acf{C. elegans} are simple organisms that are getting very popular for whole brain organization studies. The point of this problem is to reproduce the nervous system of the \ac{C. elegans}. This is done using recorded data of the input of 4 neurons and the output of 4 other neurons.

The dataset is great for our study because :
\begin{itemize}
  \item It comes from a very recent paper, and means the research is current.
  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}). In the (very) long term, a full parts of the human brain could be replaced by a very low powered chip.
  \item The paper was written by Dr. Barbulescu who happens to be one of my supervisor.
\end{itemize}

\begin{figure}[H]
  \centering
  \begin{minipage}{\columnwidth}
    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
    \hfill
    \subfloat[Test dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioTest0}}%
  \end{minipage}
  \begin{minipage}{\columnwidth}
    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain1.svg}}%
    \hfill
    \subfloat[Validation dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
  \end{minipage}
  \caption{\ac{C. elegans} dataset samples}
  \label{graph:celegans}
\end{figure}

\subsubsection{Dataset format}

The dataset contains a set of fourty set of input/output. Each input/output set contains $500ms$ of data with a time step of $0.5ms$, making a thousand points for each neurons. This data is recorded after applying current to the input neurons (PLML2, PLMR, AVBL, AVBR), and monitoring the output of four neurons (DB1, LUAL, PVR, VB1) that are known to have strong activity during  a specific behavior of the nematode known as \ac{FCM}. The simulation is done using a known model of \ac{C. elegans} connectome (complete overview of the brains connections) \cite{celegans}.

\Cref{graph:celegans} shows four sets of input/output our of the fourty that were recorded for the dataset.

The inputs for the systems are then going to be the current values for the four input neurons at every of the thousand time steps and plans to get an output as close as possible to the modeled output for all time steps.

\subsubsection{Network configuration}

The layers used to solve this problem are listed below :

\begin{itemize}
  \item An \ac{LSTM} with eight hidden states and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
  \item A time distributed dense layer with an output size of four. It is time distributed because it needs to compute every time steps and not only the last.
\end{itemize}

The \ac{LSTM} was chosen with a number of hidden states of eight because it gets the best results \cite{celegans}.

\Cref{fig:celegansModel} is a graphical representation of the model just described.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{datasets/celegansModel}
  \caption{Model used to solve the \ac{C. elegans} problem}
  \label{fig:celegansModel}
\end{figure}