### 2023-10-10 Appendices/appendixA.tex
Appendix wei2res

\fancychapter{TODO determine name}
\label{ap:a}

\section{Buffer}\label{apsec:buffer}

This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.

\section{Inverting amplifier}\label{apsec:invAmp}

This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10. Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).

A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something

\section{Solving weight to resistance}\label{apsec:wei2res}

\begin{equation}\label{eq:wei2res3}
  \begin{cases}
    w=\frac{R_f}{R_+}-\frac{R_f}{2\cdot R_f -R_+}\\
    R_-=2\cdot R_f -R_+
  \end{cases}
\end{equation}

\begin{equation}\label{eq:wei2res4}
  \begin{cases}
    w=\frac{2\cdot R_f^2-2\cdot R_+\cdot R_f}{2\cdot R_f\cdot R_+ -R_+^2}\\
    R_-=2\cdot R_f -R_+
  \end{cases}
\end{equation}

\begin{equation}\label{eq:wei2res5}
  \begin{cases}
    2\cdot R_f\cdot R_+\cdot w -R_+^2\cdot w=2\cdot R_f^2-2\cdot R_+\cdot R_f\\
    R_-=2\cdot R_f -R_+
  \end{cases}
\end{equation}

\begin{equation}\label{eq:wei2res6}
  \begin{cases}
    R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0\\
    R_-=2\cdot R_f -R_+
  \end{cases}
\end{equation}

\begin{equation}\label{eq:wei2res61}
  R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0
\end{equation}

We must now solve \cref{eq:wei2res61}.

\begin{equation}\label{eq:wei2res61}
  \Delta = 4\cdot R_f^2\cdot(w+1)^2-8\cdot w\cdot R_f^2 = 4\cdot R_f^2\cdot(w^2+1) \ge 0
\end{equation}

Giving the two solutions :

\begin{equation}\label{eq:wei2res62}
  R_{+,0}=\frac{R_f\cdot(w+1)-R_f\cdot\sqrt{w^2+1}}{w}
\end{equation}
\begin{equation}\label{eq:wei2res63}
  R_{+,1}=\frac{R_f\cdot(w+1)+R_f\cdot\sqrt{w^2+1}}{w}
\end{equation}