### 2023-10-16 2.State/main.tex
added new chapters

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The State:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{State of the art}
\label{cap:state}

\input{2.State/ai.tex}
\input{2.State/nn.tex}
\input{2.State/rnn.tex}
\input{2.State/lstm.tex}
\input{2.State/memristors.tex}
\input{2.State/crossbar.tex}
\input{2.State/model.tex}

\cleardoublepage