### 2023-10-16 2.State/rnn.tex

\section{\acs{RNN}}\label{sec:rnn}

\acp{RNN} are, as the name suggests, a type of \ac{NN} using recurrent connections. They are a \ac{NN} with at least one cycle within the structure, where outputs of the previous run is used the next one.