### 2023-10-17 2.State/ai.tex
Correction

\section{\acl{AI}}\label{sec:ai}

\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning. So what are \acp{AI} exactly.

\ac{AI} is defined as theories and stragies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules. The definition is very wide and have lots of ways to be implemented.

The most promising type of \ac{AI} is machine learning, which consists of teaching the software to make decision. It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.

\subsection{Machine learning}

In the last 10 years, the research concerning machine learning has considerably increased.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{ml.svg}
  \caption{Machine learning and its subsets}
  \label{fig:ml}
\end{figure}

\Cref{fig:ml} is a visual representation of machine learning and all the disciplines of machine learning. The thesis will focus, as the title implies, on the \aclp{RNN}.