### 2023-10-19 introduction/contributions.tex

\section{Main Contributions}
\label{sec:contributions}

In the thesis the following points will worked on in order to create, improve or clarify any of those topics :

\begin{itemize}
\item A script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
\item An \ac{LSTM} capable analog circuit.
\item Analog activation function circuit that supports \ac{tanh} and sigmoid ranges.
\item A \ac{VMM} capable analog circuit using a matrix of memristors.
\item Memory cells to store analog data for a short amount of time.
\item A python (tensorflow library) code to use piece wise functions, to reproduce in python the analog activation functions.
\item A way to run software simulation of different \acp{NN}.
\end{itemize}