### 2023-10-19 introduction/outline.tex
Forgot to rename the folders in the mains...

\section{Thesis Outline}
\label{sec:outline}

This thesis is going to tackle the work that went into creating a software simulation of an analog computer capable of running \aclp{RNN}.

All the technologies used and, thus are required to know in order to understand the thesis are detailed in the state of the art.

Next the work that was done for the thesis is goign to be thoroughly explained. Starting with the different circuit that were designed for the thesis. Then the different scripts and python code that were created, either to run the digital \acp{NN} or to generate the analog system into a netlist.