### 2023-10-19 0.Intro/abstract.tex
first draft of abstract

\begin{abstract}

  \acp{AI} are one of the main research domain in computer sciences and is used in other scientific domains. One great application of \ac{AI}, is in embedded systems. Lots of small electronic devices around us could benefit us. \textbf{Google} and \textbf{Apple} are pushing \ac{AI} in phones with their in house designed ARM processors having, very advertised, tensor cores.
  Being able to run low powered \acp{AI} is thus one of the biggest computer objective of the deceny.
  Analog computers are known to be a low powered technology that most importantly give almost instant results. An analog \ac{ASIC} could then be fabricated for specific embedded applications. This would definitely improve digital technologies in terms of time, and maybe in power consumption as well. Which are two prerequisite for embedded system use.
  For this reason this work aims at creating a functioning software simulation of analog circuit capable of running \ac{AI}. The thesis will focus on the software simulation, the first step of the work in order to have a fully working chip able to run \ac{AI} algorithms.

\end{abstract}