### 2023-10-19 0.Intro/acknowledgments.tex
thanking people and all

\pdfbookmark{Acknowledgments}{Acknowledgments}
\begin{acknowledgments}

  I would like to thank the \ac{IST} for allowing to join their double degree program, that made it possible for me to widen my computer knowledge from computer science to electronics.
  I would also like to thank \ac{INESC} to have let me join their rank, to both \ac{INESC}-Id and \ac{INESC}-MN groups to have given the necessary tools to succeed with this thesis.
  I would also like to thank all the other student working in the \ac{ASIC} group with whom I could share problems and were happy to help.
  I would lastly like to thank Dr Ruxandra Barlulescu and Dr Diogo Caetano, to have trusted with this very interresting topic despite my recent arrival in the univerty.

\end{acknowledgments}
\clearpage
\thispagestyle{empty}
\cleardoublepage