### 2023-10-24 introduction/contributions.tex
Started contibs

\section{Main Contributions}
\label{sec:contributions}

The thesis bring contributions to the following points:

\begin{itemize}
  \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
  \item An \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
  \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
  \item A \ac{VMM} capable analog circuit using a matrix of memristors.
  \item Memory cells to store analog data for a short amount of time.
  \item A python (tensorflow library) code to use piece wise functions, to reproduce in python the analog activation functions.
  \item A way to run software simulation of different \acp{NN}.
\end{itemize}