### 2023-10-25 introduction/contributions.tex
Forgot s s

\section{Main Contributions}
\label{sec:contributions}

The thesis bring contributions to the following points:

\begin{itemize}
  \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
  \item The script can generate an \ac{LSTM} capable analog circuits was created and ran in virtuoso giving encouraging results.
  \item The script can also generate an \ac{GRU} capable analog circuits was created and ran in virtuoso.
  \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
  \item I recreated a \acp{VMM} capable analog circuit using a matrix of memristors.
  \item I designed a memory cells used to store analog data for a short amount of time, they are used the \ac{LSTM} and \ac{GRU} circuits.
  \item I coded in python \cite{python} (tensorflow library\cite{tensorflow}) to create a piece wise functions to reproduce in the tensorflow the analog activation functions.
  \item I used Cadence's Virtuoso to run software simulation of different \acp{NN} capable analog circuits.
\end{itemize}