### 2023-10-26 0.Intro/abstract.tex
Changed abstract

\begin{abstract}

  In this work we propose an analog structure for memristor based recurrent modules targeting neural computing. The system is fully analog and implements a working \ac{LSTM} circuit block and a work in progress \ac{GRU} circuit block. Both of those blocks contain memristors to be used as weights in a analog \ac{VMM} capable circuit. These circuit blocks allow to run very fast computation of \aclp{RNN} of any size, in a relatively small integrated circuit. As part of the \ac{LSTM} and \ac{GRU} blocks, an analog activation function circuit was designed. This specific circuit is capable of reproducing sigmoid and \ac{tanh} like functions, with similar shapes and the same output ranges. The work also include the implementation of a memory cell used to store an analog value for a short period of time. The \ac{LSTM} block can be serialized or not with the ability to choose the level of serialization. Serializing the system allows to save onChip area at the cost of execution time. To the author's knwoledge this is the first analog implementation of the behavior of \acs{C. elegans} using the \ac{LSTM} block. Such an analog system provides ground for real time implementation of nervous systems.

\end{abstract}