### 2023-10-27 introduction/outline.tex

\section{Thesis Outline}
\label{sec:outline}

This thesis is going to tackle the work that went into creating a software simulation of an analog computer capable of running \aclp{RNN}.

In \cref{cap:state}, all the technologies used and thus required to know in order to understand the thesis are detailed in the state of the art.

Next, in \cref{cap:design}, the work that was done for the thesis is going to be thoroughly explained. Starting with the different circuit that were designed for the thesis. Then the different scripts and python code that were created, either to run the digital \acp{NN} or to generate the analog system into a netlist.

Then \cref{cap:results} contains the results, obtained from the digital inference and the analog simulation, from both the number of airline passenger and the \ac{C. elegans} datasets.

Finally, in \cref{cap:conclusions}, the conclusions and improvements that can be made to the different parts of the thesis.