### 2023-10-28 appendices/codes.tex

\fancychapter{Important codes}
\label{ap:codes}

\section{Smoothing the curve}\label{apsec:smoothFunc}

\pythonstyle
\lstinputlisting[language=python]{code/smooth.py}