### 2023-10-28 appendices/circuits.tex
New file plus working on circuits

\fancychapter{Less important circuits}
\label{ap:circuit}

\section{Buffer}\label{apsec:buffer}

This is a classic circuit that only takes one \ac{opAmp} wired as \cref{fig:buffer}. It is used to forward a voltage without interfering the node of the input voltage.

\begin{figure}[H]
\centering
\includesvg[width=.6\textwidth]{appendix/buffer}
\caption{Buffer circuit}
\label{fig:buffer}
\end{figure}

The input is $V_{in}$ and the output is $V_{out}$.

\section{Inverting amplifier}\label{apsec:invAmp}

This is the circuit that comes after the voltage multiplier (\cref{sec:models}), it is here to amplify the results of the voltage multiplier by 10. Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).