### 2023-11-02 conclusions/imag.tex
Impact environmental imag de merde

\section{Environmental and societal impact}

\subsection{Personnal environmental impact}

\subsection{Global impact}

\subsubsection{Environmental impact}

\subsubsection{Societal impact}

\subsection{Host structure's policy}