### 2023-11-30 paper/state.tex

\section{State of the art}\label{sec:state}

\subsection{Recurrent Neural Networks (RNN)}

RNNs are a family of neural networks that differentiate themselves by having feedback connections. It is often used with sequences of data \cite{rnn}, sometimes, with varying amount of input data. RNNs are used for handwriting recognition and language translation \cite{gru}.

The feedback connection is caracterized by the hidden state vector ($\symvh_t$) which serves as the output and as part of the input for the next time step, a RNN is defined as \cref{eq:rnn}.

\begin{equation}\label{eq:rnn}
  \symvh_t=f(\symvx_t,\symvh_{t-1})
\end{equation}

The simple version of the RNN is defined by \cref{eq:srnn}.

\begin{equation}\label{eq:srnn}
  \symvh_t=tanh([\symvx_t,\symvh_{t-1}]\cdot \symmw + \symvb)
\end{equation}

Where (\symmw,\symvb) are the pair of weights matrix and bias vectors. $\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.

The graphical representation of the simple RNN cell is shown in \cref{fig:rnnCell}.

\begin{figure}[t]
  \centering
  \begin{minipage}{\columnwidth}
    \subfloat[Simple \acs{RNN} cell\label{fig:rnnCell}]{\includesvg[width=\columnwidth]{rnn/rnnCell.svg}}
  \end{minipage}
  \begin{minipage}{\columnwidth}
    \subfloat[Legend\label{leg:cells}]{\includesvg[width=\columnwidth,pretex=\footnotesize]{cellsLegend.svg}}
  \end{minipage}
  \caption{}
\end{figure}

\subsection{Long Short Term Memory (LSTM)}

LSTMs are a type of RNN used to solve the vanishing gradiant problem \cite{firstLSTM}. It was improved a few times before becoming the modern LSTM \cite{improvLSTM}. The LSTM differs from a simple RNN because of its second feedback variable being the cell state ($\symvc_t$).

The LSTM contains four activated gates, that each serve its own purpose. The input gate \cref{eq:inputG} controls whether the cell state is updated, the forget gate \cref{eq:forgetG} that determines how the current cell state is affected by the old cell state, the output gate \cref{eq:outputG} that controls how much the hidden state is affected by the cell state. The candidate cell state gate \cref{eq:candCell} computes the change in the future cell state.

\begin{equation}\label{eq:inputG}
  \symvi_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_i + \symvb_i)
\end{equation}
\begin{equation}\label{eq:forgetG}
  \symvf_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_f + \symvb_f)
\end{equation}
\begin{equation}\label{eq:outputG}
  \symvo_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_o + \symvb_o)
\end{equation}
\begin{equation}\label{eq:candCell}
  \symvct_t=tanh([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_c+ \symvb_c)
\end{equation}

The next part of the LSTM consist of computing the new cell state in the following equation :

\begin{equation}\label{eq:cellS}
  \symvc_t=\symvf_t\odot \symvc_{t-1} + \symvi_t \odot \symvct_t
\end{equation}

The hidden state is then determined using the current cell state \cref{eq:hiddenS}.

\begin{equation}\label{eq:hiddenS}
  \symvh_t=\symvo_t\odot tanh(\symvc_t)
\end{equation}

Where ($\symmw_i$,$\symvb_i$), ($\symmw_f$,$\symvb_f$), ($\symmw_o$,$\symvb_o$) and ($\symmw_c$,$\symvb_c$) are the pair of weights matrixes and bias vectors for the input, forget, output and candidate cell gates respectively. $\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.

The graphical representation of an LSTM cell is found in \cref{fig:lstmCell}.

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth]{lstm/lstmCell.svg}
  \caption{\acs{LSTM} cell, adapted from \cite{wikiLSTM}\label{fig:lstmCell}}
\end{figure}

\subsection{Gated Recurrent Units (GRU)}

GRUs are another type of RNN. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.

The \ac{GRU} is very often compared to the \ac{LSTM}, being sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}. Their performance was found to be very similar in most situations \cite{gruVSlstm}, making those two types of \acp{RNN} coexistant in the modern machine learning world.

There are two versions of the \ac{GRU}, both are found on the internet, they are known as the encoder and decoder version \cite{gru}. They were originally designed to encode the message to translate and then decode in the translation. PyTorch only supports the decoder version \cite{gruPyTorch}, while the Keras library supports both \cite{gruKeras} chosen by changing an argument. This work only uses the encoder version, for that reason, it will be the only one described.

It contains an update gate \cref{eq:updateG}, a reset gate \cref{eq:resetG}, a candidate activation gate \cref{eq:candActivG}. The hidden state is then computed \cref{eq:gruHidG} using the previous results.

\begin{equation}\label{eq:updateG}
  \symvz_t= \sigma([\symvx_t,\symvh_{t-1}] \cdot \symmw_z + \symvb_z)
\end{equation}
\begin{equation}\label{eq:resetG}
  \symvr_t= \sigma([\symvx_t,\symvh_{t-1}] \cdot \symmw_r + \symvb_r)
\end{equation}
\begin{equation}\label{eq:candActivG}
  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symmw_{hh} + \symvb_h)
\end{equation}
\begin{equation}\label{eq:gruHidG}
  \symvh_t=(\vunit-\symvz_t)\odot \symvh_{t-1} + \symvz_t\odot \symvhh_t
\end{equation}

Where ($\symmw_z$,$\symvb_z$), ($\symmw_r$,$\symvb_r$),($\symmw_{hx}$,$\symmw_{hh}$,$\symvb_h$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.

A visual representation of the encoder \ac{GRU} cell is available in \cref{fig:encoderGruCell}.

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth]{gru/encoderCell.svg}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}\label{fig:encoderGruCell}}
\end{figure}

\subsection{Memristors}

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 as a missing fundamental component in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The missing component linking the four fundamental circuit variables, voltage (\symv), charge (\symq), current (\symi) and flux (\symphi). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :

\begin{equation}
  d\symphi = \symv\cdot d\symt
\end{equation}

\begin{equation}
  d\symq = \symi\cdot d\symt
\end{equation}

Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux (\symphi) and charge (\symq).  The flux in this case is not a magnetic flux and is defined as such : $ d\symphi=\symv\cdot d\symt \Rightarrow \symphi =  \int \symv \,d\symt  $.

The component stayed theoretical until 2008 when it was implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.

An extension to the memristor, reffered to as the memristive device, was theorized in 1976 \cite{memrestiveDev}. The difference between the memristor and the memristive devices is its internal behavior. Memristive devices are commonly referred to as memristors as well.

\begin{figure}[t]
  \centering
  \includesvg[width=0.65\columnwidth,pretex=\tiny]{memristor/memristor}
  \caption{Fundamental passive components, adapted from \cite{memWiki}}
  \label{fig:fundComp}
\end{figure}

A memristor is used for its ability to change its internal resistance based on the current that flowed through it.

\subsection{Memristors Crossbar Arrays}

Setting memristors in a crossbar array allows to perform analog Vector Matrix Multiplication (VMM), also called Multiply and Accumulate. \Cref{fig:crossbar} shows what a typical crossbar array looks like.

\begin{figure}[b]
  \centering
  \includesvg[width=.6\columnwidth,pretex=\small]{crossbar/crossbar}
  \caption{Crossbar array schematics, inspired from \cite{xbarFigures}}
  \label{fig:crossbar}
\end{figure}

The circuit uses physical properties of electrical systems to perform analog computation. The following part will focus only on the circuit node in \cref{fig:crossNode}.

\begin{figure}[t]
  \centering
  \includesvg[wdith=0.4\columnwidth,pretex=\scriptsize]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $k^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

A voltage is applied on the $k^{th}$ line, and because every column is virtually grounded, the voltage applied to the memristor, with resistance $R_i$, is $V_i$. By applying Ohm's law, we know that the current flowing into the memristor ($\symi_{k}$) is bound by the following equation :

\begin{equation}
  \symv_k = \symR\cdot \symi_{k} \Rightarrow \symi_{k} = \symv_k\cdot (\frac{1}{\symR_k})= \symv_k\cdot \symG_k
\end{equation}
With $ \symG_k$ being the conductance of memristor, defined as $ \symG_k=\frac{1}{\symR_k}$.

This line then joins the column where a current of $\symi_{j,k-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
\begin{equation}
  \symi_{j,k} = \symi_{j,k-1}+\symi_{k} = \symi_{j,k-1} + \symv_k\cdot \symG_k
\end{equation}
Unfolding the equations will give the current at the bottom of the column, for example, the current at the bottom of the first column in \cref{fig:crossbar} is :
\begin{equation}
  \symi_1= \symG_1\cdot \symv_1 +  \symG_2\cdot \symv_2 +  \symG_3\cdot \symv_3
\end{equation}
With $\symG_1$, $\symG_2$ and $\symG_3$ being the conductance of the 3 memristors in the first column.