### 2023-12-13 conclusions/imag.tex
Started imag envirnoment

\section{Environmental and societal impact}

\subsection{Personnal environmental impact}

\subsection{Global impact}

The subject of this thesis being purely theoretical, running simulations only, the environmental and societal impact will focus on the system if it were to become real. This assumes that the entire system is working and is being fabricated for embedded use. The final product can be used in both comsumer and business oriented product.

\subsubsection{Environmental impact}

\subsubsection{Societal impact}

\subsection{Host structure's policy}