### 2023-12-13 state/main.tex

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The State:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{State of the art}
\label{cap:state}

\input{state/ai.tex}
\input{state/nn.tex}
\input{state/rnn.tex}
\input{state/lstm.tex}
\input{state/gru.tex}
\input{state/memristors.tex}
\input{state/crossbar.tex}
\input{state/model.tex}

\cleardoublepage